---
title: EDA
output: md_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = FALSE)
```

```{r}
library(magrittr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(purrr)
library(readxl)

source("../R/dm.R")

df <- readr::read_csv("../data/data_v1.0_country-year.csv") %>%
  select(-year)

df_acq <- readr::read_csv("../data/data_v1.0_country-year-mode_acq.csv", 
                          na = c("n.a.", "n.a", "")) %>%
  filter(article != "No provision") %>%
  select(-year) %>% 
  mutate(cntry_mode_id = paste(iso3, mode_id, sep = "_"))

df_acq_proc <- dm_procedure(df_acq)

df_loss <- readr::read_csv("../data/data_v1.0_country-year-mode_loss.csv", na = c("n.a.", "n.a", "")) %>% 
  filter(article != "No provision") %>%
  select(-year) %>%
  mutate(cntry_mode_id = paste(iso3, mode_id, sep = "_"))

df_loss_proc <- dm_procedure(df_loss)

df_long <- df %>%
  tidyr::pivot_longer(names_to = "mode_id", values_to = "values", -c(iso3, country, region)) %>% 
  tidyr::separate(mode_id,
                  into = c("mode_id", "type"),
                  sep = "([(_)])") %>%
  mutate(par_mode_id = substr(mode_id, 1, 3),
         acq = grepl("A", mode_id),
         loss = grepl("L", mode_id),
         bin = "bin" == type,
         cat = "cat" == type,
         cntry_mode_id = paste(iso3, mode_id, sep = "_"))

df_long %>% 
  readr::write_csv("../data/cit_long.csv")
```

## Number of unique modes

```{r}
acq_mode <- df_long %>% filter(acq) %>% distinct(par_mode_id) %>% pull()
n_acq <- length(acq_mode)

loss_mode <- df_long %>% filter(loss) %>% distinct(par_mode_id) %>% pull()
n_loss <- length(loss_mode)

n_cntry <- df %>% distinct(country) %>% pull() %>% length()
```

* Number **acquisition** variables: `r n_acq`
* **Acquisition** variables: `r acq_mode`

* Number **loss** variables: `r n_loss`
* **Loss** variables: `r loss_mode`

* Number of countries: `r n_cntry`

## Countries with 99 in A06 mode

```{r}
df_long %>%
  filter(bin) %>%
  filter(acq) %>%
  filter(values == 99) %>% 
  distinct(country)
```

## Most common acq modes

```{r}
acq_mode_common <- df_long %>%
  filter(bin) %>%
  filter(acq) %>%
  filter(values == 1) %>%
  distinct(par_mode_id, country, region) %>% 
  group_by(par_mode_id, region) %>% 
  tally() %>% 
  arrange(desc(n))

ggplot(acq_mode_common, aes(x = reorder(par_mode_id, -n), y = n, fill = reorder(region, n))) +
  geom_bar(stat="identity") +
  labs(y = "N",
       x = "Mode",
       title = "Number of countries where a citizen can gain citizenship by mode") +
  ylim(0, n_cntry)
```

```{r}
loss_mode_common <- df_long %>%
  filter(bin) %>%
  filter(loss) %>%
  filter(values == 1) %>%
  distinct(par_mode_id, country, region) %>% 
  group_by(par_mode_id, region) %>% 
  tally() %>% 
  arrange(desc(n))

ggplot(loss_mode_common, aes(x = reorder(par_mode_id, -n), y = n, fill = reorder(region, n))) +
  geom_bar(stat="identity") +
  labs(y = "N",
       x = "Mode",
       title = "Number of countries where a citizen can lose citizenship by mode") +
  ylim(0, n_cntry)
```

## Acquiring citizenship by country

```{r}
acq_top_10 <- df_long %>%
    filter(bin) %>%
    filter(acq) %>%
    filter(values == 1) %>% 
    group_by(country, region) %>% 
    tally() %>% 
    arrange(n) %>%
    ungroup()

ggplot(acq_top_10 %>% 
    slice(1:10), 
       aes(x = reorder(country, n), y = n)) +
  geom_bar(stat="identity") +
  labs(y = "N",
       x = "Country",
       title = "Top ten countries with the fewest ways to acquire citizenship") +
   theme(axis.text.x = element_text(angle = 65, vjust = 0.5, hjust=1))
```

```{r}
ggplot(acq_top_10,
       aes(x = n, fill = region, color = region)) +
  geom_density(alpha = .5) +
  labs(y = "Density",
       x = "Number",
       title = "Number of ways to acquire citizenship by country") +
   theme(axis.text.x = element_text(angle = 65, hjust=1))
```

* Median number of ways to acquire citizenship: `r median(acq_top_10$n)`

## Procedure of acquisition

```{r}
ggplot(df_acq_proc %>% 
         group_by(procedure) %>% 
         tally(), 
       aes(x = reorder(procedure, -n), y = n)) +
  geom_bar(stat="identity") +
  labs(y = "N",
       x = "Procedure",
       title = "Procedure to acquire citizenship",
       caption = "There may be more than one procedure per acquisition mode") +
   theme(axis.text.x = element_text(angle = 65, hjust=1))

```

## Losing citizenship by country

```{r}
loss_top_10 <- df_long %>%
    filter(bin) %>%
    filter(loss) %>%
    filter(values == 1) %>% 
    group_by(country, region) %>% 
    tally() %>% 
    arrange(desc(n)) %>%
    ungroup()

ggplot(loss_top_10 %>% 
    slice(1:10), 
       aes(x = reorder(country, -n), y = n)) +
  geom_bar(stat="identity") +
  labs(y = "N",
       x = "Country",
       title = "Top ten countries with the most ways to lose citizenship") +
   theme(axis.text.x = element_text(angle = 65, vjust = 0.5, hjust=1))
```

```{r}
ggplot(loss_top_10,
       aes(x = n, fill = region, color = region)) +
  geom_density(alpha = .5) +
  labs(y = "Density",
       x = "Number",
       title = "Number of ways to lose citizenship by country") +
   theme(axis.text.x = element_text(angle = 65, hjust=1))
```

* Median number of ways to lose citizenship: `r median(loss_top_10$n)`

## Procedure of acquisition

```{r}
ggplot(df_loss_proc %>% 
         group_by(procedure) %>% 
         tally(), 
       aes(x = reorder(procedure, -n), y = n)) +
  geom_bar(stat="identity") +
  labs(y = "N",
       x = "Procedure",
       title = "Procedure to loss citizenship",
       caption = "There may be more than one procedure per loss mode") +
   theme(axis.text.x = element_text(angle = 65, hjust=1)) + 
  theme_minimal()

```


## Visualizing similarity in countries aquisitions spacially

```{r}
ggplot(df %>%
         select(country, region, contains("bin")) %>%
         select(country, region, starts_with("A")) %>%
         pivot_longer(values_to = "values", 
                      names_to = "mode_id", 
                      -c(country, region)) %>%
        mutate(mode_id = gsub("_cat", "", mode_id),
               mode_id = gsub("_bin", "", mode_id)) %>% 
         filter(values != "99") %>%
         dplyr::group_by(mode_id) %>% 
         slice(1:14),
       aes(x = mode_id, y = values, fill = region, color = region, group = factor(country))) +
  geom_area(position = "identity") +
  geom_point() +
  facet_wrap(vars(country)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        axis.text.y = element_blank()) +
  labs(x = "",
       y = "") +
  coord_polar()
```

### Comparing colonized countries with their colonizer

```{r}
ggplot(df %>%
         select(country, iso3, region, contains("bin")) %>%
         select(country, iso3, region, starts_with("A")) %>%
         pivot_longer(values_to = "values", 
                      names_to = "mode_id", 
                      -c(country, iso3, region)) %>%
        mutate(mode_id = gsub("_cat", "", mode_id),
               mode_id = gsub("_bin", "", mode_id)) %>% 
         filter(values != "99") %>%
         dplyr::group_by(mode_id) %>% 
         filter(iso3 %in% c("GBR", "AUS", "NZL", "IND", "ZAF", "USA")),
       aes(x = mode_id, y = values, fill = region, color = region, group = factor(country))) +
  geom_area(position = "identity", alpha = .5) +
  geom_point() +
  facet_wrap(vars(country)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        axis.text.y = element_blank()) +
  labs(x = "",
       y = "") +
  coord_polar()
```

## Initial Observations

* All countries share at one way to gain citizenship (A01)
* There isn't one universal way citizens can lose citizenship
* L13 doesn't apply to any countries in Africa, Oceania, or the Americas
* The average European country has more ways to acquire citizenship than the average country from another region of the world
* There is a more uniform distribution across how many ways there are to lose citizenship
* The procedure of acquisition can differ from country to country. For example, A01 isn't always automatic Attribution
* There may be some overlap in acquisition laws depending on if the country was colonized

