---
title: EDA
output: md_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = FALSE)
```

```{r}
library(tidyverse)

source("../R/dm.R")
source("../R/question_coding.R")

lu_loc <- read_csv("../data/lu_code.csv")
xwalk <- read_csv("../data/xwalk_region.csv")
flow <- read_csv("../data/flow.csv")

df <- readr::read_csv("../data/data_v1.0_country-year.csv") %>%
  select(-year)

df_acq <- readr::read_csv("../data/data_v1.0_country-year-mode_acq.csv", 
                          na = c("n.a.", "n.a", "")) %>%
  filter(article != "No provision") %>%
  select(-year) %>% 
  mutate(cntry_mode_id = paste(iso3, mode_id, sep = "_"))

df_loss <- readr::read_csv("../data/data_v1.0_country-year-mode_loss.csv", na = c("n.a.", "n.a", "")) %>% 
  filter(article != "No provision") %>%
  select(-year) %>%
  mutate(cntry_mode_id = paste(iso3, mode_id, sep = "_"))

df_acq_proc <- dm_procedure(df_acq)
df_loss_proc <- dm_procedure(df_loss)

df_long <- df %>%
  tidyr::pivot_longer(names_to = "mode_id", values_to = "values", -c(iso3, country, region)) %>% 
  tidyr::separate(mode_id,
                  into = c("mode_id", "type"),
                  sep = "([(_)])") %>%
  mutate(par_mode_id = substr(mode_id, 1, 3),
         acq = grepl("A", mode_id),
         loss = grepl("L", mode_id),
         bin = "bin" == type,
         cat = "cat" == type,
         yrs = "yrs" == type,
         cntry_mode_id = paste(iso3, mode_id, sep = "_"))

bin <- df_long %>%
  filter(acq & bin & values == 1) %>%
  select(-c(acq, type, loss, cat, yrs, bin, par_mode_id, cntry_mode_id)) %>%
  filter(mode_id %in% unique(modes$mode_id))

df_long <- df_long %>%
  filter(acq & c(cat | yrs) & values > 0) %>%
 select(-c(acq, type, loss, par_mode_id, cntry_mode_id))

subregion_code <- xwalk %>% 
  distinct(subregion, subregion_code)

codes <- subregion_code %>% 
  distinct(subregion_code) %>% 
  pull(subregion_code)

modes <- dm.combine_questions(df_long, df_acq) %>%
  inner_join(df %>% select(iso3, country))

```


```{r write_data}

z <- flow %>%
  filter(dest_loc %in% codes) %>%
  filter(orig_loc %in% codes) %>%
  select(-c(dest_notes, dest_dtype)) %>%
  dplyr::group_by(dest_loc, orig_loc) %>%
  summarize(value = sum(value)) %>%
  rename(DEST = dest_loc,
         ORIG = orig_loc) %>%
  filter(DEST != ORIG) %>%
  ungroup() %>%
  mutate(id = 1:n(),
         value_bins = case_when(value < 1000 ~ "< 1000",
                                value >= 1000 & value < 10000 ~ "< 10000",
                                value >= 10000 & value < 100000 ~ "< 100000",
                                value >= 100000 & value < 1000000 ~ "< 1000000",
                                value >= 1000000 ~ "1000000+"))

z %>%
  write_csv("../svelte-development/public/data/region_flows.csv")

z %>%
  select(-value) %>%
  pivot_longer(names_to = "PLACE", values_to = "CODE", -c(id)) %>% 
  left_join(centroids) %>% 
  write_csv("../svelte-development/public/data/region_flows_geo.csv")
  

b <- flow %>%
  filter(dest_loc %in% xwalk$loc_code) %>%
  filter(orig_loc %in% xwalk$loc_code) %>%
  dplyr::group_by(dest_loc) %>%
  summarize(value = sum(value)) %>%
  rename(loc_code = dest_loc) %>%
  # filter(DEST) %>%
  ungroup() %>%
  inner_join(xwalk) %>%
  select(iso3, region, value) %>% 
  rename(country = region) %>% 
  write.csv("../svelte-development/public/data/country_flows.csv")

total_migrants <- flow %>%
  filter(dest_loc %in% codes) %>%
  filter(orig_loc %in% codes) %>%
  group_by(orig_loc) %>%
  summarize(TOTAL_MIGRANTS = sum(value)) %>%
  inner_join(subregion_code, by = c("orig_loc" = "subregion_code")) %>%
  mutate(TOTAL_MIGRANTS_BINS = case_when(TOTAL_MIGRANTS < 1000000 ~ "< 1 million",
                              TOTAL_MIGRANTS >= 1000000 & TOTAL_MIGRANTS < 2000000 ~ "< 2 million",
                              TOTAL_MIGRANTS >= 2000000 & TOTAL_MIGRANTS < 5000000 ~ "< 5 million",
                              TOTAL_MIGRANTS > 5000000 ~ "> 5 million"))

centroid <- geojsonR::FROM_GeoJson("../svelte-development/public/data/mapData/region-centroid.geojson")

centroid$features <- lapply(centroid$features, function(x) {
  x2 <- x$properties
  
  ttl <- total_migrants %>%
    filter(orig_loc == x2$CODE)
  
  x2$TOTAL_MIGRANTS <- ttl %>%
    pull(TOTAL_MIGRANTS)
  
  x2$TOTAL_MIGRANTS_BIN <- ttl %>%
    pull(TOTAL_MIGRANTS_BINS)
  
  return(x2)
  
})

library(sf)
meuse_sf = sf::st_as_sf(centroid, coords = c("x", "y"), crs = 28992, agr = "constant")
st_write(meuse, "meuse.geojson")

"../svelte-development/public/data/mapData/region-centroid.geojson"

region_geo <- dm.region_geojson(pth = "../data/shape/world-administrative-boundaries.shp",
                                xwalk,
                                write_pth = "../data/region_geo.geojson")

modes %>%
  write_csv("../data/modes_acq.csv")

df_long %>%
 readr::write_csv("../data/cit_long.csv")

df_long %>%
  filter(mode_id %in% unique(modes$mode_id)) %>% 
  readr::write_csv("../data/allCountriesAllPaths.csv")


df_long %>%
  filter(mode_id == "A06a" & cat) %>%
  select(iso3, country, mode_id, values) %>%
  mutate(time = case_when(values == 1 ~ "Fewer than 5 years",
                          values == 2 ~ "Exactly 5 years",
                          values == 3 ~ "Between 5 years and 10 years",
                          values == 4 ~ "More than 10 years")) %>%
  inner_join(modes %>%
               filter(mode_id == "A06a") %>% 
               select(mode_id)) %>%
  readr::write_csv("../svelte-development/public/data/modeA06a.csv")
  

```

```{r}
ggplot(z %>% filter(value <= 100000), aes(x = value)) +
  geom_histogram(bins = 100, color = "white")

```

## Number of unique modes

```{r}
acq_mode <- df_long %>% filter(acq) %>% distinct(par_mode_id) %>% pull()
n_acq <- length(acq_mode)

loss_mode <- df_long %>% filter(loss) %>% distinct(par_mode_id) %>% pull()
n_loss <- length(loss_mode)

n_cntry <- df %>% distinct(country) %>% pull() %>% length()
```

* Number **acquisition** variables: `r n_acq`
* **Acquisition** variables: `r acq_mode`

* Number **loss** variables: `r n_loss`
* **Loss** variables: `r loss_mode`

* Number of countries: `r n_cntry`

## Countries with 99 in A06 mode

```{r}
df_long %>%
  filter(bin) %>%
  filter(acq) %>%
  filter(values == 99) %>% 
  distinct(country)
```

## Most common acq modes

```{r}
acq_mode_common <- df_long %>%
  filter(bin) %>%
  filter(acq) %>%
  filter(values == 1) %>%
  distinct(par_mode_id, country, region) %>% 
  group_by(par_mode_id, region) %>% 
  tally() %>% 
  arrange(desc(n))

ggplot(acq_mode_common, aes(x = reorder(par_mode_id, -n), y = n, fill = reorder(region, n))) +
  geom_bar(stat="identity") +
  labs(y = "N",
       x = "Mode",
       title = "Number of countries where a citizen can gain citizenship by mode") +
  ylim(0, n_cntry)
```

```{r}
loss_mode_common <- df_long %>%
  filter(bin) %>%
  filter(loss) %>%
  filter(values == 1) %>%
  distinct(par_mode_id, country, region) %>% 
  group_by(par_mode_id, region) %>% 
  tally() %>% 
  arrange(desc(n))

ggplot(loss_mode_common, aes(x = reorder(par_mode_id, -n), y = n, fill = reorder(region, n))) +
  geom_bar(stat="identity") +
  labs(y = "N",
       x = "Mode",
       title = "Number of countries where a citizen can lose citizenship by mode") +
  ylim(0, n_cntry)
```

## Acquiring citizenship by country

```{r}
acq_top_10 <- df_long %>%
    filter(bin) %>%
    filter(acq) %>%
    filter(values == 1) %>% 
    group_by(country, region) %>% 
    tally() %>% 
    arrange(n) %>%
    ungroup()

ggplot(acq_top_10 %>% 
    slice(1:10), 
       aes(x = reorder(country, n), y = n)) +
  geom_bar(stat="identity") +
  labs(y = "N",
       x = "Country",
       title = "Top ten countries with the fewest ways to acquire citizenship") +
   theme(axis.text.x = element_text(angle = 65, vjust = 0.5, hjust=1))
```

```{r}
ggplot(acq_top_10,
       aes(x = n, fill = region, color = region)) +
  geom_density(alpha = .5) +
  labs(y = "Density",
       x = "Number",
       title = "Number of ways to acquire citizenship by country") +
   theme(axis.text.x = element_text(angle = 65, hjust=1))
```

* Median number of ways to acquire citizenship: `r median(acq_top_10$n)`

## Procedure of acquisition

```{r}
ggplot(df_acq_proc %>% 
         group_by(procedure) %>% 
         tally(), 
       aes(x = reorder(procedure, -n), y = n)) +
  geom_bar(stat="identity") +
  labs(y = "N",
       x = "Procedure",
       title = "Procedure to acquire citizenship",
       caption = "There may be more than one procedure per acquisition mode") +
   theme(axis.text.x = element_text(angle = 65, hjust=1))

```

## Losing citizenship by country

```{r}
loss_top_10 <- df_long %>%
    filter(bin) %>%
    filter(loss) %>%
    filter(values == 1) %>% 
    group_by(country, region) %>% 
    tally() %>% 
    arrange(desc(n)) %>%
    ungroup()

ggplot(loss_top_10 %>% 
    slice(1:10), 
       aes(x = reorder(country, -n), y = n)) +
  geom_bar(stat="identity") +
  labs(y = "N",
       x = "Country",
       title = "Top ten countries with the most ways to lose citizenship") +
   theme(axis.text.x = element_text(angle = 65, vjust = 0.5, hjust=1))
```

```{r}
ggplot(loss_top_10,
       aes(x = n, fill = region, color = region)) +
  geom_density(alpha = .5) +
  labs(y = "Density",
       x = "Number",
       title = "Number of ways to lose citizenship by country") +
   theme(axis.text.x = element_text(angle = 65, hjust=1))
```

* Median number of ways to lose citizenship: `r median(loss_top_10$n)`

## Procedure of acquisition

```{r}
ggplot(df_loss_proc %>% 
         group_by(procedure) %>% 
         tally(), 
       aes(x = reorder(procedure, -n), y = n)) +
  geom_bar(stat="identity") +
  labs(y = "N",
       x = "Procedure",
       title = "Procedure to loss citizenship",
       caption = "There may be more than one procedure per loss mode") +
   theme(axis.text.x = element_text(angle = 65, hjust=1)) + 
  theme_minimal()

```


## Visualizing similarity in countries aquisitions spacially

```{r}
ggplot(df %>%
         select(country, region, contains("bin")) %>%
         select(country, region, starts_with("A")) %>%
         pivot_longer(values_to = "values", 
                      names_to = "mode_id", 
                      -c(country, region)) %>%
        mutate(mode_id = gsub("_cat", "", mode_id),
               mode_id = gsub("_bin", "", mode_id)) %>% 
         filter(values != "99") %>%
         dplyr::group_by(mode_id) %>% 
         slice(1:14),
       aes(x = mode_id, y = values, fill = region, color = region, group = factor(country))) +
  geom_area(position = "identity") +
  geom_point() +
  facet_wrap(vars(country)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        axis.text.y = element_blank()) +
  labs(x = "",
       y = "") +
  coord_polar()
```

### Comparing colonized countries with their colonizer

```{r}
ggplot(df %>%
         select(country, iso3, region, contains("bin")) %>%
         select(country, iso3, region, starts_with("A")) %>%
         pivot_longer(values_to = "values", 
                      names_to = "mode_id", 
                      -c(country, iso3, region)) %>%
        mutate(mode_id = gsub("_cat", "", mode_id),
               mode_id = gsub("_bin", "", mode_id)) %>% 
         filter(values != "99") %>%
         dplyr::group_by(mode_id) %>% 
         filter(iso3 %in% c("GBR", "AUS", "NZL", "IND", "ZAF", "USA")),
       aes(x = mode_id, y = values, fill = region, color = region, group = factor(country))) +
  geom_area(position = "identity", alpha = .5) +
  geom_point() +
  facet_wrap(vars(country)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        axis.text.y = element_blank()) +
  labs(x = "",
       y = "") +
  coord_polar()
```

## Initial Observations

* All countries share at one way to gain citizenship (A01)
* There isn't one universal way citizens can lose citizenship
* L13 doesn't apply to any countries in Africa, Oceania, or the Americas
* The average European country has more ways to acquire citizenship than the average country from another region of the world
* There is a more uniform distribution across how many ways there are to lose citizenship
* The procedure of acquisition can differ from country to country. For example, A01 isn't always automatic Attribution
* There may be some overlap in acquisition laws depending on if the country was colonized
