---
title: EDA
output: md_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = FALSE)
```

```{r}
library(magrittr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(purrr)

df <- readr::read_csv("../data/data_v1.0_country-year.csv")

df_long <- df %>%
  tidyr::pivot_longer(names_to = "sub_code", values_to = "values", -c(iso3, country, year, region)) %>% 
  mutate(bin = grepl("bin", sub_code),
         acq = grepl("A", sub_code),
         loss = grepl("L", sub_code),
         code = substr(sub_code, 1, 3)
         )
```

## Number of unique codes

```{r}
acq_codes <- df_long %>% filter(acq) %>% distinct(code) %>% pull()
n_acq <- length(acq_codes)

loss_codes <- df_long %>% filter(loss) %>% distinct(code) %>% pull()
n_loss <- length(loss_codes)

n_cntry <- df %>% distinct(country) %>% pull() %>% length()
```

* Number **acquisition** variables: `r n_acq`
* **Acquisition** variables: `r acq_codes`

* Number **loss** variables: `r n_loss`
* **Loss** variables: `r loss_codes`

* Number of countries: `r n_cntry`

## Countries with 99 in A06 code

```{r}
df_long %>%
  filter(bin) %>%
  filter(acq) %>%
  filter(values == 99) %>% 
  distinct(country)
```

## Most common acq codes

```{r}
acq_code_common <- df_long %>%
  filter(bin) %>%
  filter(acq) %>%
  filter(values == 1) %>%
  distinct(code, country, region) %>% 
  group_by(code, region) %>% 
  tally() %>% 
  arrange(desc(n))

ggplot(acq_code_common, aes(x = reorder(code, -n), y = n, fill = reorder(region, n))) +
  geom_bar(stat="identity") +
  labs(y = "N",
       x = "Code",
       title = "Number of countries where a citizen can gain citizenship by code") +
  ylim(0, n_cntry)
```

```{r}
loss_code_common <- df_long %>%
  filter(bin) %>%
  filter(loss) %>%
  filter(values == 1) %>%
  distinct(code, country, region) %>% 
  group_by(code, region) %>% 
  tally() %>% 
  arrange(desc(n))

ggplot(loss_code_common, aes(x = reorder(code, -n), y = n, fill = reorder(region, n))) +
  geom_bar(stat="identity") +
  labs(y = "N",
       x = "Code",
       title = "Number of countries where a citizen can lose citizenship by code") +
  ylim(0, n_cntry)
```

## Initial Observations

* All countries share at one way to gain citizenship (A01)
* There isn't one universal way citizens can lose citizenship
* L13 doesn't apply to any countries in Africa, Oceania, or the Americas
